ScPy: A SuperCollider Extension for Performing Numerical Computation via Embedded Python

### 2016-08-25 doc/paper.tex
supposedly vertical rules in tables are ugly. add some references too

\usepackage{fancyhdr}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{minted}
\usepackage[backend=bibtex]{biblatex}

@@

    SuperCollider, a language for sound synthesis and algorithmic composition of audio, supports a
    wide range of synthesis, effect and analysis algorithms. However, available operations are
    limited to those implemented explicitly as Unit Generators (UGens). Since UGens are written in
    C/C++ and loaded as plugins during the SuperCollider server boot process, it is impossible to
    live code UGens, which limits the user to creating sound as a composition of existing UGens
    during a performance.  Many of the vector operations required for efficiently creating complex
    audio effects are notably missing or tedious to use. To overcome this, we present ScPy, a UGen
    which embeds Python within SuperCollider to enable the use of the optimized vector operations
    provided by the NumPy library.

@@

Although a number of open-source projects which interface other languages with SuperCollider exist,
they are primarily SuperCollider clients, which have features to control synths and send other
automation messages. One notable exception is OctaveSC, which embeds GNU Octave within
SuperCollider. However, it is designed primarily for performing operations on control rate data
arrays and its author notes performance issues which prevent it from being used on audio rate data.

@@

Although designing a domain-specific language (DSL) for this task was briefly considered, we
eventually decided to use Python for a number of reasons. First, Python includes an excellent C API
which makes embedding it within other languages trivial in comparison to many other options.
Second, Python's syntax is easily readable and concise -- it clearly adheres to our goals. Finally,
Python has a massive number of community built libraries for performant scientific and numerical
computing. With a few small exceptions, these libraries contained every one of the fundamental
spectral operations we could think of.

@@

When \autoref{lst:boilerplate} is run, an anonymous synth is created which will take in audio
from an external source, perform a STFT, process the spectral data with ScPy, perform an inverse
STFT, then finally output the resulting audio. Lines 8-9 contain Python code defining the
processing function where operations are to be inserted. Further examples will replace only those
lines. Line 16 contains the Python code to convert the spectrum into a NumPy array, process it,
then write back the processed version.

@@

\begin{table}[H]
    \caption{Type correspondence between SuperCollider and Python}
    \begin{center}
        \begin{tabular}{| l | l |}
            \hline
            SuperCollider data type & Python data type \\ \hline
            \texttt{Float} & \texttt{float} \\ \hline
            \texttt{UGen} (subclass of) & \texttt{ndarray} of \texttt{float} \\ \hline
            \texttt{Buffer} & \texttt{ndarray} of \texttt{float} \\ \hline
            \texttt{FFT} & \texttt{ndarray} of \texttt{complex} \\ \hline
            \texttt{Array} & \texttt{list} \\ \hline
        \end{tabular}
    \end{center}
\end{table}

@@

\printbibliography

>>>>

\usepackage{fancyhdr}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{minted}
\usepackage{biblatex}
\usepackage{booktabs}

@@

    SuperCollider~\cite{wilson2011supercollider}, a language for sound synthesis and algorithmic
    composition of audio, supports a wide range of synthesis, effect and analysis algorithms.
    However, available operations are limited to those implemented explicitly as Unit Generators
    (UGens). Since UGens are written in C/C++ and loaded as plugins during the SuperCollider server
    boot process, it is impossible to live code UGens, which limits the user to creating sound as a
    composition of existing UGens during a performance.  Many of the vector operations required for
    efficiently creating complex audio effects are notably missing or tedious to use. To overcome
    this, we present ScPy, a UGen which embeds Python within SuperCollider to enable the use of the
    optimized vector operations provided by the NumPy library.

@@

Although a number of open-source projects which interface other languages with SuperCollider exist
\cite{systemsinterfacingwithsc}~\cite{magnusson2011ixi}~\cite{orlarey2009faust}, they are mostly
SuperCollider clients, which have features to control synths and send other automation messages.
One notable exception is OctaveSC~\cite{octavesc}, which embeds GNU Octave within SuperCollider.
However, it is designed primarily for performing operations on control rate data arrays and its
author notes performance issues which prevent it from being used on audio rate data.

@@

Although designing a domain-specific language (DSL)~\cite{van2000domain} for this task was briefly
considered, we eventually decided to use Python for a number of reasons. First, Python includes an
excellent C API which makes embedding it within other languages trivial in comparison to many other
options.  Second, Python's syntax is easily readable and concise --- it clearly adheres to our
goals. Finally, Python has a massive number of community built libraries for performant scientific
and numerical computing. With a few small exceptions, these libraries contained every one of the
fundamental spectral operations we could think of.

@@

When \autoref{lst:boilerplate} is run, an anonymous synth is created which will take in audio
from an external source, perform a STFT, process the spectral data with ScPy, perform an inverse
STFT, then finally output the resulting audio. Lines 8--9 contain Python code defining the
processing function where operations are to be inserted. Further examples will replace only those
lines. Line 16 contains the Python code to convert the spectrum into a NumPy array, process it,
then write back the processed version.

@@

\begin{table}[H]
    \caption{Type correspondence between SuperCollider and Python}
    \begin{center}
        \begin{tabular}{cc}
            \toprule
            SuperCollider type & Python type \\
            \midrule
            \texttt{Float} & \texttt{float} \\
            (subclass of) \texttt{UGen} & \texttt{ndarray} of \texttt{float} \\
            \texttt{Buffer} & \texttt{ndarray} of \texttt{float} \\
            \texttt{FFT} & \texttt{ndarray} of \texttt{complex} \\
            \texttt{Array} & \texttt{list} \\
            \bottomrule
        \end{tabular}
    \end{center}
\end{table}

@@

\printbibliography{}
